Methods for Large Scale Learning to Rank

### 2014-04-22 report/Chapters/LTRMethods.tex
smoothrank uitgelegd :)

\section{SmoothRank}
SmoothRank \cite{Chapelle2010} is a listwise ranking method that directly optimises an information retrieval evaluation measure by smoothing, that is, approximating the rank position. In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure. The smoothing function used in SmoothRank based on the softmax activation function \cite{Bridle1990} that is often used in neural networks, which is formally defined as shown in \ref{eq:softmax_algorithm}.
\begin{equation}
p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{m}e^{f_j/\sigma}}
\label{eq:softmax_algorithm}
\end{equation}

% TODO: 1) Smoothing framework uitleggen en toepassen op nDCG 2) sigma trade-off behandelen 3)Nonlinear conjugate gradient descent uitleggen 4) startin point en regularization uitleggen 5) Annealing process uitleggen %

>>>>

\section{SmoothRank}
SmoothRank \cite{Chapelle2010} is a listwise ranking method that directly optimises an information retrieval evaluation measure by smoothing, that is, approximating the rank position. In this section I will use \ac{nDCG} for illustration of SmoothRank, but the same procedure can also be applied to \ac{MAP} or any other information retrieval measure. The smoothing function used in SmoothRank based on the softmax activation function \cite{Bridle1990} that is often used in neural networks, which is formally defined as shown in Equation \ref{eq:softmax_algorithm}.
\begin{equation}
p_i = \frac{e^{f_i/\sigma}}{\sum\nolimits_{j=1}^{m}e^{f_j/\sigma}}
\label{eq:softmax_algorithm}
\end{equation}

where $\sigma$ is the smoothing parameter. Chapelle et al \cite{Chapelle2010} apply the softmax activation function to the \ac{nDCG} formula and thereby introduce a soft version of indicator variable $h_{i,j}$ as formulated in Equation \ref{eq:soft_ndcg}.

\begin{equation}
h_{i,j} = e^{-\frac{(f(x_i))-(f(x_{d(j)}))^2}{\sigma}}\Big/\sum\nolimits_{k=1}^{m}e^{-\frac{(f(x_k))-f(x_{d(j)})^2}{\sigma}}
\label{eq:soft_ndcg}
\end{equation}

It can be shown that the derivative of the smoothed \ac{nDCG} version shown in Equation \ref{eq:soft_ndcg} and the smoothed versions of other \ac{IR} metrics can be calculated in $\mathcal{O}(m^2)$, which enable fast gradient descent optimisation. The optimisation step in SmoothRank uses the nonlinear conjugate gradient method  with Polak-Ribiere update \cite{Shewchuk1994}, which is a type of gradient descent method. The optimisation method is prone to local optima, which is alleviated by adding a good starting point and regulariser.\\

The starting point in SmoothRank is set to either the solution of a simple linear regression, or alternatively to the solution of Rank\acs{SVM}. Since this starting point is expected to already be a good solution, a regulariser term is added to the SmoothRank objective function to prevent the solution from deviating to much from the starting point. The regularised smooth objective function is formulated as $\lambda||w-w_0||^2$ where $\lambda$ is a hyperparameter tuned on the validation set, and $w_0$ is the starting point solution.\\

The choice of the smoothing parameter $\sigma$ in Equation \ref{eq:soft_ndcg} is important, because a too small value makes the function more non-smooth and therefore harder to optimise, while a too large value results in a optimisation function that substantially differs from the optimal rankings. SmoothRank uses an annealing method where the optimisation procedure starts with a large $\sigma$ and iteratively reduces it by dividing it by 2 at each step. Algorithm \ref{alg:smoothrank} shows the algorithm that summarises all steps of the SmoothRank method.
